---
title: "Treatment mapping"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ironer)
# ?treatment_map
# View(treatment_map)
```

When possible, treatment regimens are mapped to standardized coding through a a master treatment mapping file.
This master list is created by identifying not yet mapped treatments and then personnel researching the drug class, which is confirmed and QC'd by medical monitors when in question.
The treatment mapping is continuously updated.

The treatment mapping data set can be accessed by installing this package and then
submitting `View(treatment_map)`. The treatment mapping is also shown here
for convenience. Please submit corrections or additions as an [issue](https://github.com/pcctc/ironer/issues) in this
repository, or contact the PCCTC data science team at <zzpdl_pcctc_data_science@mskcc.org>.



| Variable             | Meaning                        | 
| :--------------------|:---------------------------------
| `treatment`          | Treatment as submitted to the study |
| `treatment_lower`    | Treatment as submitted to the study lower case |
| `treatment_recode`   | Assigned treatment standardized coding |
| `treatment_category` | Assigned treatment category |
| `is_taxel`           | Indicates if treatment is a `taxel` (one of Cabazitaxel, Docetaxel, Paclitaxel, Tesetaxel) |
| `is_nha`             | Indicates if treatment is a novel hormone agent (one of Abiraterone, Enzalutamide, Darolutamide, Apalutamide) |


```{r dictionary-reactable, echo=FALSE}
treatment_map %>% 
   reactable::reactable(
    filterable = TRUE,
    searchable = TRUE,
    resizable = TRUE,
    #groupBy = "treatment_category",
    defaultPageSize = 500
    #paginateSubRows = FALSE,
    #columns = list(
    #  treatment_category = reactable::colDef(minWidth = 200)
    #)
  )
```
