---
title: "Comprehensive Ironman Treatment Tutorial"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, warning=FALSE, echo=FALSE, message=FALSE}
# load packages
library(tidyverse)
library(lubridate)
library(labelled)
library(stringr)
library(janitor)
library(ironer)
library(reactable)
library(dsverse)
```


```{r, echo=FALSE}
# import list of data frames exported from medidata
lst_medidata <- lst_blind_medidata

```

# Overview of treatment data

Treatment data is collected in two independent forms:\

1. Cancer treatments (`ca_cm`)\

2. Physician questionnaires ([multiple versions implemented across protocols](https://pcctc.github.io/ironer/articles/data-documentation-medidata.html#physician-questionnaires))\

Data capture differs between the forms; as the forms are independent, treatments may not align across the forms. The Data Management team monitors this and issues queries to improve alignment. Your Ironman research question should inform which source of treatment data to use, and if it is useful to combine the two sources of treatment information.

## Cancer treatments (`ca_cm`)

Cancer treatments represent treatments that occurred both prior to study enrollment and on study. Data is collected as log lines, with one row per treatment. Cancer treatments could be used for any time-to-event analyses where exact start and stop date of treatments is required.

Each entry includes treatment name, start date, stop date, modification reason, discontinuation reason, and indication. Indication was added June 2024 when the prior and on study treatment forms were consolidated to a single form.

The `ca_cm` table represents the raw cancer treatment data collected in medidata. The `ca_cm_derived` table additionally contains higher level treatment categories, standardized treatment coding, and flags for taxels and NHAs (novel hormonal agents).

### Additional considerations for analysis:

1. It could be assumed that all patients on anti-androgen are simultaneously on ADT, 
even if ADT is not reported. Note this may vary globally.

2. It could be assumed that a patient is continuing ADT if it was ever marked as "so far".

3. It could be assumed that patients with mHSPC are on ADT, even if not marked. 
This can be cross checked against treatments.

4. If a systemic treatment is listed in "so far" but "starting" is blank, the visit associated with "so far" treatment could be imputed as their initial on study treatment.

5. Consider ignoring external beam radiation therapy in treatment regimen. (“We often will have patients with mHSPC get radiation to a painful lesion or something along with starting systemic therapy.)

6. There is data on orchiectomies in the `prsurg` and `surg` tables that can be incorporated.


### Gather minimal data from cancer treatments form

```{r}
df_treatments_all <- lst_medidata$ca_cm_derived |> 
  select(
    subject,
    treatment = trt_treatment_recode,
    dt_treatment_start = exstdat_int,
    dt_treatment_end = exendat_int,
    treatment_category = trt_treatment_category
  ) |> 
  # remove blank treatments
  filter(!is.na(treatment))|>
  # sort by treatment start date within subject
  arrange(subject, dt_treatment_start)

```

Here is an example of what the data looks like for a single patient. Dates have been shifted to protect the patient's identity.

```{r, echo=FALSE, results=TRUE}
df_treatments_all |>
  filter(subject == "1")|>
  select(-subject)|>
  mutate(dt_treatment_start = dt_treatment_start,
         dt_treatment_end = dt_treatment_end) |>
  reactable()
  
```

### Treatments and treatment categories

Depending on your research question, you may want to exclude treatments or treatment categories that are not relevant for your analysis. The table below displays all treatments and their categorization.

```{r}
df_treatments_all |>
  count(treatment_category, treatment) |> 
  reactable()
```

Be aware that the ADT (Androgen Deprivation Therapy) treatment classification includes
LHRH agonists, LHRH antagonists, and anti-androgens. Depending on your research question,
you may want to implement a different treatment classification that more narrowly
defines ADTs of interest.

### Removing ineligible treatments

Occasionally treatments that are not specifically cancer treatments are entered in medidata. The following code excludes treatment categories that are not specific to prostate cancer. This
list should be regularly assessed to ensure it is up to date and accurate as new
data gets entered.

```{r}
# vector of treatment categories that are not specifically prostate cancer treatments
# and could be excluded from analysis of prostate cancer specific treatments
trt_category_exclude <- c(
  "Steroid", "Antidiabetic", "Antilipemic", "Nutraceuticals", "Bone-Modifying Agent",
  "Bisphosphonates", "Nutraceutical", "Nutraceutical/Placebo", "Placebo", "COX-2 Inhibitor"
  )

# create a new data frame that represents prostate cancer (pc) treatments only 
df_treatments_pc <- df_treatments_all |>
  # exclude these drug classes
  filter(!treatment_category %in% trt_category_exclude)
```


### Collapsing consecutive treatments

Occasionally treatments are entered on different log lines but with no break
between the treatment end date and the next treatment start date. For example, consider
this subject history which shows Bicalutamide continuing between rows 5 and 6,
and then again between rows 7 and 8. 

```{r, echo=FALSE}
tibble::tribble(
     ~row,     ~treatment, ~dt_treatment_start, ~dt_treatment_end,
       1L, "Bicalutamide",        "2001-04-09",      "2003-04-01",
       2L, "Bicalutamide",        "2008-06-01",      "2008-09-01",
       3L, "Bicalutamide",        "2009-06-24",      "2009-12-01",
       4L, "Bicalutamide",        "2010-10-08",      "2011-06-01",
       5L, "Bicalutamide",        "2012-09-01",      "2013-11-26",
       6L, "Bicalutamide",        "2013-11-26",      "2014-12-01",
       7L, "Bicalutamide",        "2015-01-01",      "2016-07-26",
       8L, "Bicalutamide",        "2016-07-27",      "2017-04-04"
     )
```

For analysis purposes, this could be considered a single treatment of Bicalutamide 
represented by 6 rows instead of 8. 

```{r, echo=FALSE}
tibble::tribble(
     ~row,     ~treatment, ~dt_treatment_start, ~dt_treatment_end,
       1L, "Bicalutamide",        "2001-04-09",      "2003-04-01",
       2L, "Bicalutamide",        "2008-06-01",      "2008-09-01",
       3L, "Bicalutamide",        "2009-06-24",      "2009-12-01",
       4L, "Bicalutamide",        "2010-10-08",      "2011-06-01",
       5L, "Bicalutamide",        "2012-09-01",      "2014-12-01",
       6L, "Bicalutamide",        "2015-01-01",      "2017-04-04"
     )
```

This code will collapse consecutive or overlapping records
of the same treatment. 

```{r}
df_treatments_pc_collapsed <- df_treatments_pc |> 
   arrange(
     subject, 
     treatment,
     dt_treatment_start, 
     desc(dt_treatment_end)
     ) |>
  mutate(
    # examine each treatment within a subject
    .by = c(subject, treatment),
    # create a flag that represents truly unique lines of treatment
    flag_unique_line = case_when(
      # the first record represents a unique line
      row_number() == 1 ~ TRUE,
      # if treatment start date is after the previous treatment end date plus one,
      # then we assume it represents a new line for that treatment.
      dt_treatment_start > lag(dt_treatment_end) + 1 ~ TRUE,
      # otherwise, if the treatment start date occurs before the previous treatment 
      # end date plus one, this is not a unique line
      dt_treatment_start <= lag(dt_treatment_end) + 1 ~ FALSE,
      # for instance of missing start or end date, the default value will assume a new new line
      .default = TRUE),
    # create a grouping variable that represents the total number of unique lines
    group = cumsum(flag_unique_line),
  ) |> 
  # # examine each treatment grouping within a subject
  # for overlapping entries for a drug, take the earlier start date and later end date
  mutate(
    .by = c(subject, treatment, group),
    dt_treatment_start_group = first(dt_treatment_start, na_rm = TRUE),
    dt_treatment_end_group = last(dt_treatment_end, na_rm = TRUE)
  ) |> 
  distinct(
    subject, 
    treatment, 
    dt_treatment_start = dt_treatment_start_group, 
    dt_treatment_end = dt_treatment_end_group, 
    treatment_category
  ) |> 
  arrange(
    subject, dt_treatment_start
  )

```

### Assigning Line of Therapy (LOT)

The treatments form captures individual treatments, but does not capture if treatments
are administered simultaneously or line of therapy (LOT) for that treatment.

The line of the therapy for a given treatment can be approximately derived using
the [`ironer::assign_lot`](https://pcctc.github.io/ironer/reference/assign_lot.html) function.
Treatments are considered to be in the same LOT if their start dates are within 30 days. Treatment end dates are not considered in defining LOTs.

**Considerations:**

-   ADT usage

    - ADT is not consistently captured, and the documentation or usage of ADT varies globally.
    
    - About 50% of mHSPC patients have ADT documented in the database.
    
    - You could consider excluding ADT when creating lines of therapy, and assume ADT usage 
    even if not documented.

-   The code below considers all treatments initiated within 90 days of enrollment as per protocol patients must be enrolled within 90 days of starting systemic treatment. This can be modified as needed.

-   The function can be used to assign line of therapy based on either the treatment category or the specific treatment itself. 


```{r}
# identify on study treatments
df_treatments_onstudy <- df_treatments_pc_collapsed |>
  # include consent date in order to exclude treatments >90 days before enrollment
  left_join(lst_medidata$ic |> select(subject, cnstdate_int), by = "subject") |>
  mutate(
    # day of treatment relative to consent date
    trt_day_rel_consent = dt_treatment_start - cnstdate_int,
    # consider treatment to be on study if within 90 days of consent date
    trt_on_study = if_else(trt_day_rel_consent >= -90, TRUE, FALSE)
  ) |> 
  # retain only on study treatments
  filter(trt_on_study)
```

Identify first line of therapy on study.

```{r}
# assign LOT while on study, by treatment category
# be patient, this takes a bit to run
# the result is the same dimension as the input data, i.e., one row per treatment
df_lot_trt <- df_treatments_onstudy |> 
  arrange(subject, dt_treatment_start) |>
  group_by(subject) |>
  group_modify(~ assign_lot(
    .x, treatment_category, dt_treatment_start, dt_treatment_end
  )) |>
  ungroup() 

# modify the result to be one row per line of therapy
# additionally retain all treatments in that line of therapy 
df_lot <- df_lot_trt |> 
  select(
    subject, 
    cnstdate_int,
    treatment,
    lot_on_study = lot, 
    regimen_on_study = regimen, 
    dt_lot_start, 
    dt_lot_last_obs
  ) |> 
  # occasionally, identical treatments are consecutive and estimated to be part of 
  # the same LOT. distinct() here reduces the identical and consecutive treatments
  # to one row
  distinct() |>
  # create an indicator for observed treatments, to be used in pivot
  mutate(trt_ind = 1) |> 
  # create a single variable that represents all treatment in LOT
  arrange(subject, lot_on_study, treatment) |> 
  group_by(subject, lot_on_study) |> 
  mutate(trt_on_study = str_c(treatment, collapse = ", "), .after = regimen_on_study) |> 
  ungroup() |> 
  # this makes treatment columns show in alphabetical order in the pivot step
  arrange(treatment) |> 
  # wrangle the data from long to wide to create 1 row per LOT, with indicators
  # for each individual treatment
  pivot_wider(
    # creates column names from the pivot
    names_from = treatment,
    # observed treatments are given a value of 1
    values_from = trt_ind,
    # treatments not observed are filled with a value of 0
    values_fill = 0
    ) |> 
  arrange(subject, lot_on_study, dt_lot_start)
```

Here is an example of what the LOT data looks like for a single patient. 
Dates have been shifted to protect the patient's identity.

```{r, echo=FALSE, results=TRUE}

df_lot |>   
  filter(subject=="1")|>   
  select(-subject)|>   
  mutate(dt_lot_start = dt_lot_start,          
         dt_lot_last_obs = dt_lot_last_obs) |>
  reactable()
```

## Treatments from Physician Questionnaires (PQ)

The Physician Questionnaire (PQ) should be completed by the treating physician at
baseline and at every 3-month visit. Each questionnaire asks what treatments the
patient has received so far and which treatments the patient is starting. There are
currently five versions of the PQ that must be combined for analysis (PQs were amended
at various protocol amendments). Each version is a separate table in the database 
with different variable names. The following table shows the questions and responses 
for each version of the PQ.

PQ treatments are useful to assess the cross-sectional association between treatment
and other PQ or PROM measures that are linked to a visit. There are no specific dates
for treatment in the PQ, so it is challenging to use them for time-to-event analyses. 


The date the PQ was entered in the edc is in the field `record_date` (though
this does not necessarily correspond to the date the PQ was completed by the 
physician). Each PQ is associated with a visit (like `Baseline`, `Month 3`, etc),
and those visit dates could be brought in from the `cyc_date` form.

+------------------+-------------------+------------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------+
| PQ version       | Table in Database | Questions asked                                                                                                                          | Responses                                                                        |
+==================+===================+==========================================================================================================================================+==================================================================================+
| Versions 1 and 2 | pq                | Please list this patient’s CURRENT / CHANGE IN TREATMENT(S)                                                                              | Open text field                                                                  |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   | Are you planning on continuing androgen deprivation therapy for this patient?                                                            | Yes/No                                                                           |
+------------------+-------------------+------------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------+
| Version 3        | new_pq            | What systemic treatments has your patient previously received for his prostate cancer? Choose all that apply.                            | -   Androgen deprivation therapy (surgical castration, LH-RH agonist/antagonist) |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   | What treatment is your patient starting? May select more than one if in combination                                                      |     -   option for first question only                                           |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Enzalutamide                                                                 |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Abiraterone acetate                                                          |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Docetaxel chemotherapy                                                       |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Radium-223 dichloride                                                        |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Sipuleucel-T                                                                 |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Cabazitaxel chemotherapy                                                     |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Bicalutamide, or other anti-androgen                                         |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Clinical Trial                                                               |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Other \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_                                         |
+------------------+-------------------+------------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------+
| Version 4        | new_pq_v4         | What systemic treatment(s) has your patient received so far and which one(s) is he starting now? In both columns, choose all that apply. | -   Abiraterone                                                                  |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Androgen deprivation (LHRH, surgical)                                        |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Apalutamide                                                                  |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Bicalutamide or other anti-androgen                                          |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Cabazitaxel                                                                  |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Clinical trial                                                               |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Docetaxel                                                                    |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Enzalutamide                                                                 |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   External beam radiation therapy for metastatic disease                       |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Radium-223                                                                   |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Sipuleucel-T                                                                 |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Other cancer-directed therapy (e.g., mitoxantrone)                           |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Symptom-oriented therapy only (e.g., analgesics)                             |
+------------------+-------------------+------------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------+
| Version 5        | new_pq_v5         | What systemic treatment(s) has your patient received so far and which one(s) is he starting now? In both columns, choose all that apply.  | -   Abiraterone                                                                  |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Androgen deprivation (LHRH, surgical)                                        |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Apalutamide                                                                  |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Bicalutamide or other anti-androgen                                          |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Cabazitaxel                                                                  |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Clinical trial                                                               |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Docetaxel                                                                    |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Enzalutamide                                                                 |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   External beam radiation therapy for metastatic disease                       |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Radium-223                                                                   |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Sipuleucel-T                                                                 |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Other cancer-directed therapy (e.g., mitoxantrone)                           |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Symptom-oriented therapy only (e.g., analgesics)                             |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Darolutamide                                                                 |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Niraparib                                                                    |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Olaparib                                                                     |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Rucaparib                                                                    |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Talazoparib                                                                  |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   Pembrolizumab                                                                |
|                  |                   |                                                                                                                                          |                                                                                  |
|                  |                   |                                                                                                                                          | -   177Lu-PSMA-617                                                               |
+------------------+-------------------+------------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------+



### PQ version 1 and 2

Protocol versions 1 and 2 use the same version of the physician questionnaire.
Due to the open text nature, it is challenging to standardize treatments.

```{r}
# use variable naming conventions that align with other pq aggregations
df_pq1 <- lst_medidata$pq |> 
  mutate(pq_version = "pq1") |> 
  select(pq_version, subject, instance_name, treatment = pq1, continuing_adt = pq2) 
```


### PQ version 3

For PQ3, pull data in two separate steps: one for so far treatments and one for
starting treatments. Then combine.

```{r}
# Retrieve so far treatments
df_pq3_sofar <- lst_medidata$new_pq |>
  select(subject, instance_name, matches("^newpq_4[a-i]$"), newpq_4iab, -contains("raw")) |> 
  # standardize treatment name through label
  labelled::set_variable_labels(
    newpq_4a = "ADT",
    newpq_4c = "Abiraterone",
    newpq_4d = "Docetaxel",
    newpq_4e = "Radium-223",
    newpq_4g = "Cabazitaxel",
    newpq_4h = "Anti-androgen",
    newpq_4i = "Clinical trial", # to standardize name across other PQs
    subject = NULL,
    instance_name = NULL,
    newpq_4iab = "sofar_Other"
  ) |> 
  # convert label to variable name
  sjlabelled::label_to_colnames() |> 
  # create a set of new columns that contain the character string
  # representing treatment continued so far
  mutate(
    across(
      .cols = -c(subject, instance_name, sofar_Other), 
      .names = "sofar_{.col}", 
      .fns = \(x) case_when(
        x == 1 ~ var_label(x),
        .default = NA
    ))
  ) |> 
  select(subject, instance_name, contains("sofar")) |> 
  relocate(sofar_Other, .after = last_col())

# Retrieve starting treatments
df_pq3_starting <- lst_medidata$new_pq |>
  select(subject, instance_name, matches("^newpq_5[a-h]"), newpq_5ia, -contains("raw")) |> 
  # standardize treatment name through label
  labelled::set_variable_labels(
    newpq_5b = "Abiraterone",
    newpq_5c = "Docetaxel",
    newpq_5d = "Radium-223",
    newpq_5f = "Cabazitaxel",
    newpq_5g = "Anti-androgen",
    newpq_5h = "Clinical trial",  # to standardize name across other PQs
    subject = NULL,
    instance_name = NULL,
    newpq_5ia = "starting_Other",
  ) |> 
  # convert label to variable name
  sjlabelled::label_to_colnames() |> 
  # create a set of new columns that contain the character string
  # representing treatment starting
  mutate(
    across(
      .cols = -c(subject, instance_name, starting_Other), 
      .names = "starting_{.col}", 
      .fns = \(x) case_when(
          x == 1 ~ var_label(x),
        .default = NA
    ))
  ) |> 
  select(subject, instance_name, contains("starting")) |> 
  relocate(starting_Other, .after = last_col())

# Merge so far and starting treatments
df_pq3 <- df_pq3_sofar |> 
  left_join(df_pq3_starting, by = c("subject", "instance_name"))

```

### PQ version 4

```{r}
df_pq4 <- lst_medidata$new_pq_v4 |>
  select(subject, instance_name, matches("^newpq14[a-m]_v4"), -contains(c("std"))) |> 
  labelled::set_variable_labels(
    newpq14b_v4 = "ADT",
    newpq14d_v4 = "Anti-androgen",
    newpq14i_v4 = "EBRT",
    newpq14m_v4 = "Other",
    subject = NULL,
    instance_name = NULL
  ) |> 
  sjlabelled::label_to_colnames() |> 
  mutate(
    # create a set of columns for so far treatments
    across(
      .cols = -c(subject, instance_name), 
      .names = "sofar_{.col}", 
      .fns = \(x) case_when(
          x == "So far" ~ cur_column(),
        .default = NA
      )),
    # create a set of columns for starting treatments
    across(
      .cols = -c(subject, instance_name, contains("sofar")), 
      .names = "starting_{.col}", 
      .fns = \(x) case_when(
          x == "Starting now" ~ cur_column(),
          .default = NA
      )),
  ) |> 
  select(subject, instance_name, starts_with("sofar"), starts_with("starting"))


```

### PQ version 5

```{r}
df_pq5 <- lst_medidata$new_pq_v5 |>
  select(subject, instance_name, matches("^newpq14[a-u,]_v5"), -contains(c("std", "14n"))) |> 
  labelled::set_variable_labels(
    newpq14b_v5 = "ADT",
    newpq14d_v5 = "Anti-androgen",
    newpq14i_v5 = "EBRT",
    newpq14m_v5 = "Other",
    subject = NULL,
    instance_name = NULL
  ) |> 
  sjlabelled::label_to_colnames() |> 
  mutate(
    # create a set of columns for so far treatments
    across(
      .cols = -c(subject, instance_name), 
      .names = "sofar_{.col}", 
      .fns = \(x) case_when(
          x == "So far" ~ cur_column(),
        .default = NA
      )),
    # create a set of columns for starting treatments
    across(
      .cols = -c(subject, instance_name, contains("sofar")), 
      .names = "starting_{.col}", 
      .fns = \(x) case_when(
          x == "Starting now" ~ cur_column(),
          .default = NA
      )),
  ) |> 
  select(subject, instance_name, starts_with("sofar"), starts_with("starting"))
```

### Combine data from all PQs 

This creates a wide data set combining PQs versions 3, 4 and 5 with a set of 
columns for treatments continuing so far and treatment that are starting.


```{r}
# stack the three questionnaires 
df_pq_345 <- bind_rows(
    "pq3" = df_pq3,
    "pq4" = df_pq4,
    "pq5" = df_pq5,
    # create an identifier displaying questionnaire source
    .id = "pq_version"
  ) |> 
  # clean variable names 
  janitor::clean_names() |> 
  # create a single column concatenating all so far treatments
  unite(
    starts_with("sofar_"), 
    col = "sofar_trt", 
    sep = ", ", 
    remove = FALSE, 
    na.rm = TRUE
    ) |>
  # create a single column concatenating all starting treatments
  unite(
    starts_with("starting_"), 
    col = "starting_trt", 
    sep = ", ", 
    remove = FALSE, 
    na.rm = TRUE
    ) 
```

Identify the date associated with visits.

```{r}
df_visit_date <- lst_medidata$cyc_date |> 
  select(subject, instance_name, visit_date) |> 
  # remove observations with missing date
  filter(!is.na(visit_date)) |> 
  # in case of duplicate entries, retain distinct rows
  distinct()
```

Create a data frame with information from all available PQs. 

```{r}
df_pq_long <- df_pq_345 |> 
  select(-c(sofar_trt, starting_trt)) |> 
  pivot_longer(
    -c(pq_version, subject, instance_name),
    names_to = c("status", "pq_coded_trt"),
    # separate column names by first underscore
    names_pattern = "^([^_]+)_(.*)",
    values_to = "treatment",
    values_drop_na = TRUE
  ) |> 
  bind_rows(df_pq1) |> 
  # join in dates associated with visit
  left_join(df_visit_date, by = c("subject", "instance_name")) |> 
  relocate(visit_date, .after = instance_name) |> 
  arrange(subject, instance_name) |> 
# in case of duplicate entries caused by the pivot, retain distinct rows
  distinct()


```



