---
title: "PROMS data documentation"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ironer)
# ?dictionary_proms
# View(dictionary_proms)
```



# About PROMS

PROMS is a self-report questionnaire for the IRONMAN study. 
The data deliveries are a work in progress to most efficiently meet the needs of all involved partners.
The survey consists of multiple instruments. Some instruments also have multiple versions.


# PROMS dictionary

All survey instruments are contained in a single data file; each section shown corresponds to a survey instrument.
Click on the triangle to expand the survey instrument and see invidual questions.


```{r }
library(ironer)
#?dictionary_proms
```

```{r dictionary-reactable, echo=FALSE}
dictionary_proms %>% 
   reactable::reactable(
    filterable = TRUE,
    searchable = TRUE,
    resizable = TRUE,
    groupBy = "instrument",
    defaultPageSize = 15,
    #paginateSubRows = FALSE,
    columns = list(
      instrument = reactable::colDef(minWidth = 200)
    )
  )
```

# More details

## Race/ethnicity

Demographic questions were asked across two different instruments (`irondemog` and `irondemog_v3`).
Due to the common usage of race/ethnicity, these questions were combined from the two instruments for analysis convenience.
These are noted in the `irondemog combined` instrument, as well as in the medidata `subject` table.

-   `race.26.x` (a select all question, where `x` indicates the numeric select all option)

    -   `race.26.1` coalesces values from `irondemog.26.1` and `irondemog_v3.26.1`

    -   `race.26.2` coalesces values from `irondemog.26.2` and `irondemog_v3.26.2`

    -   etc.

-   `ethnicity.27`, a multiple choice question.

    -   coalesced from `irondemog.27` and `irondemog_v3.27`

## Height/weight

Survey respondents were asked height and weight in a two part question.
First, they enter the numeric value of height and weight, and second the enter the unit of measurement (inches/cm or pounds/kg).
These measurements have been kept in separate variables according to the the unit indicated.
Note that these are also in two different instruments, and should be coalesced for analysis.

-   `irondemog.3.1`, `irondemog_v3.3.1`: height in inches

-   `irondemog.3.2`, `irondemog_v3.3.2`: height in cm

-   `irondemog.5.1`, `irondemog_v3.5.1`: weight in pounds

-   `irondemog.5.2`, `irondemog_v3.5.2`: weight in kg


# FAQ
