---
title: "Medidata data documentation"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r }
library(ironer)
```

# About Medidata

Medidata is the electronic data capture system used for clinical data in IRONMAN.
The data deliveries are a work in progress to most efficiently meet the needs of all involved partners,
and is subject to change. 

# Medidata dictionary

The clinical data from medidata consists of multiple tables; each section shown
corresponds to a different data table. Click on the triangle to expand the 
data table and see individual fields. Only the first 10 tables are shown,
click on page numbers to see more tables or search in the search bar for a
field of interest.


```{r setup}
library(ironer)
# ?dictionary_medidata
# View(dictionary_medidata)
```

```{r dictionary-reactable, echo=FALSE}
dictionary_medidata %>% 
   reactable::reactable(
    filterable = TRUE,
    searchable = TRUE,
    resizable = TRUE,
    groupBy = "Table",
    defaultPageSize = 10,
    #paginateSubRows = FALSE,
    columns = list(
      Table = reactable::colDef(minWidth = 200)
    )
  )
```



# More details

## Treatments

See the tables

**`priorrx`** (medical history/prior treatments)

-   This table ideally contains prostate cancer treatments from patient's medical history, before informed consent was signed for the IRONMAN study.

**`treatment`** (cancer treatment)

-   This table contains prostate cancer treatment ongoing at time of informed consent or treatments started while on study.

While these are the completion guidelines for entry, sometimes site may enter in the wrong form so both tables should be referenced to get full view of cancer treatment.
Each of these forms in the EDC (electronic data capture) are log line forms; `record_position` is the order of entry in the form, but does not represent anything beyond that (i.e., does not represent therapy order).

## Treatment mapping (derived variables)

When possible, treatment regimens are mapped to standardized coding through a a master treatment mapping file.
This master list is created by identifying not yet mapped treatments and then personnel researching the drug class, which is confirmed and QC'd by medical monitors when in question.
The treatment mapping is continuously updated.

Standardized treatment mapping is contained in `priorrx_derived` and `treatment_derived`.
These two tables contain all information in the original `priorrx` and `treatment` tables, and additionally provide:

-   `priorrxreg_lower`, `treatment_lower`: entry in lower case

-   `priorrx_treatment_recode`, `trt_treatment_recode`: assigned standardized coding

-   `priorrx_treatment_category`, `trt_treatment_category`: assigned treatment category

-   `priorrx_is_taxel`, `trt_is_taxel`: treatment is one of Cabazitaxel, Docetaxel, Paclitaxel, Tesetaxel

-   `priorrx_is_nha`, `trt_is_nha` : treatment is a novel hormonal agent (one of Abiraterone, Enzalutamide, Darolutamide, Apalutamide)

## Gleason scores  (derived variables)

There are 6 computed / derived Gleason scores.
In general, gleason scores consist of

-   gleason 1 (max value 5)

-   gleason 2 (max value 5)

-   gleason sum (max value 10)

However, sites occasionally enter the Gleason sum in either of the first Gleason scores.
In the cleaned derived Gleason scores

-   if both gleason 1 and gleason 2 are less than or equal to five, the total gleason score is the sum of the two

-   if the first gleason score is greater than five, the total gleason score is taken to be the first gleason score

In the instance that a table has multiple gleason scores per subject (e.g., on study biopsy), the largest gleason score sum is carried forward to the subject table.

| Item | Table                 | Variable                   | Description                                                                                                               |
|-----------|-----------|-----------|----------------------------------------|
| 1    | `mhdiagbx`, `subject` | `gleason_bx_mh`            | medical history                                                                                                           |
| 2    | `prpros`, `subject`   | `gleason_rp_prpros`        | pre-study prior prostatectomy pathology                                                                                   |
| 3    | `bx_prmi`, `subject`  | `gleason_bx_prmi`          | on study biopsy                                                                                                           |
| 4    | `rp_prmi`, `subject`  | `gleason_rp_prmi`          | on study prostatectomy pathology                                                                                          |
| 5    | `subject`             | `gleason_rp_derived`       | prostatectomy pathology, utilizing first on-study results if available, and then pre-study if on-study results are absent |
| 6    | `subject`             | `gleason_combined_derived` | coalesced gleason sum, with pathological gleason prioritized over biopsy gleason (on study and medical history)           |

## Baseline Metastatic  (derived variables)

The designation of a participant as baseline metastatic looks across seven different tables.
In the `subject` table:

-   `is_metastatic_baseline = 1` if at least one of the seven tables indicates presence of metastatic disease

-   `origin_metastatic_baseline` indicates what want indicated the metastatic status across the seven tables

-   `date_metastatic_baseline` is the earliest date of metastatic flag recorded across the seven tables

| Item | Table                               | General Variable(s)                                                                                  | Date Variable                                                     | Description                                                                                                                                                                                         |
|------------|------------|------------|------------|---------------------------|
| 1    | `mhclinstg`                         | `clinstg_mstage_u_XXX_std`                                                                           | `clinstg_date_int`                                                | Medical History/Clinical Stage (TNM) at Diagnosis: looks for M stage M1, M1a, M1b, M1c                                                                                                              |
| 2    | `prpros`                            | `rpdtl_mstage_u_XXX_std`                                                                             | `prprosdt_int`                                                    | Medical History/Prior Prostatectomy: looks for M stage M1, M1a, M1b, M1c                                                                                                                            |
| 3    | `mhdiagbx`                          | `diagbx_site, diagbx_histology`                                                                      | `diagbx_date_int`                                                 | Medical History/Diagnostic Biopsy: excludes sites listed as `benign` or `negative` in histology; excludes sites containing `Prostate`,`Pelvic lymph`, `Pelvic LN`, `Unknown`, `Pelvis`, `prostatic` |
| 4    | `cs`                                | `csskull`, `csthor`, `csspine`,`cspelvis`, `csext`, `csliver`, `cslung`, `csdn`, `csoth`, `csother2` | `csdat_int`                                                       | Metastatic Disease Sites (only from `Baseline 0` folder): looks for `Yes` in any of specified fields                                                                                                |
| 5    | `new_pq` , `new_pq_v4`, `new_pq_v5` | `new_pq::newpq_2[bcde]`; `new_pq_v4::newpq12[bcde]_v4` ; `new_pq_v5::newpq12[bcde]_v5`               | `newpq_date_int` (only protocol 3; represents date QRE completed) | New Treatment Physician Questionnaire (only from `Baseline 0` folder)                                                                                                                               |
| 6    | `mhdx`                              |                                                                                                      | `mhspcdat_int`, `mcrpcdat_int`                                    | Disease state; if dates occur prior to baseline window (`cnstdate_int` + 30 days), then considered baseline metastatic                                                                                             |
| 7    | `mhaddbx`                              | `addbx_site`, `addbx_histology`                                                                   | `addbxdate_int`                 | Additional biopsies; if dates occur prior to baseline window (`cnstdate_int` + 30 days), then considered baseline metastatic; if multiple biopsies present, take earliest date

# FAQ

1. **Are genetic testing results available?**

While IRONMAN collects genetic testing reports, the information requires manual 
review to be abstracted. We are developing a CRF better capture and deliver detailed
genetic information.
