---
title: "Line of therapy algorithm"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

The line of therapy (lot) algorithm **approximately** assigns line of therapy based on the 
logic described below. 

During a manual review process, we have identified situations where
results of the algorithm do not align with the line of therapy assessment of those with domain 
expertise.

We are making this algorithm available in the hopes that it can provide a rough approximation,
but please note it is not gold standard truth. PCCTC is working towards 
electronic data base updates to more easily and accurately collect line of therapy information.
The current version of the algorithm is reasonably sufficient to extract first line of therapy.



## Data preparation

1.  Identify the data of interest. While the intent of the `priorrx` table is to capture
treatments prior to study start and the `treatment` is to capture on study treatments,
this does not always happen. A good practice is to stack both tables to create a
full treatment history and filter on dates when relevant.

 * Overall line of therapy: stack `priorrx_derived` and `treatment_derived` tables to create a full listing of prior and current treatments.
 
 * On study line of therapy: either use `treatment_derived` tables or stack `priorrx_derived` and `treatment_derived` and filter on treatments that occur after the consent date.
 
 * Disease state line of therapy: researchers may be interested in line of therapy after a specific
 disease state diagnosis, like mHSPC. If so, stack `priorrx_derived` and `treatment_derived` and filter on treatments
 that occur after the disease state diagnosis.
 
2.  Remove rows with `priorrx_treatment_category`/`trt_treatment_category` equal to any of "ADT", "Bisphosphonates", "Steroid", as these are generally not considered a standalone line of therapy.

3.  If no start date is listed, remove the row from consideration; future data cuts will hopefully be queried and have more information available.


## Line of therapy algorithm

The `assign_lot()` does this internally.

1.  Sort the table by treatment start date, and assign the first line (1L) as the first entry, along with entries that occur within a 30 day window of this date as a combo therapy.

2.  End line 1 (or subsequent lines) when a treatment stop date is listed for all therapies in that line.
    If no end date is entered, use information from step 8 if a new LOT is entered.

3.  Any new treatments listed after 30 days of an LOT start date become a new LOT.
    Any treatments listed within 30 days of this new LOT start are marked as a combo therapy for that LOT.

4.  Repeat until all LOTs are assigned for each patient.

## Example implementation 

This is a demonstration for situation 1: "Overall line of therapy." Note
that this takes a few minutes to complete.

```{r}
library(tidyverse)
library(ironer)
```


```{r, eval = FALSE, echo = TRUE}
# stack prior treatments with on study treatments
# rename variables to match between two data sets
treatment_prep <- medidata$priorrx_derived %>% 
  select(
    subject, treatment = priorrxreg, 
    exstdat_int = priorrxstdt_int, 
    exendat_int = priorrxenddt_int, 
    treatment_recode = priorrx_treatment_recode, 
    treatment_category = priorrx_treatment_category
  ) %>%
  # create variable to indicate origin of data
  mutate(source = "priorrx") %>% 
  # stack on study treatments
  bind_rows(
    medidata$treatment_derived %>% 
      select(
        subject, treatment, exstdat_int, exendat_int, 
        treatment_recode = trt_treatment_recode, treatment_category = trt_treatment_category
      ) %>% 
      # create variable to indicate origin of data
      mutate(source = "treatment")
  ) %>%
  # remove entries with missing start date
  filter(!is.na(exstdat_int)) %>% 
  # remove treatments not counted as a line of therapy
  filter(!trt_treatment_category %in% c("Steroid", "Bone-Modifying Agent", "Bisphosphonates", "Antidiabetic")) |> 
  # keep relevant fields
 select(subject, exstdat_int, exendat_int, treatment_category, treatment_recode, source)
```

Here is what the prepared data looks like for a single subject:

```{r, eval=TRUE, echo=FALSE}
tibble::tribble(
     ~subject, ~exstdat_int, ~exendat_int,      ~treatment_category,        ~treatment_recode,     ~source,
 "170-01-004", "2006-12-15", "2007-06-15",                    "ADT",           "Bicalutamide",   "priorrx",
 "170-01-004", "2011-02-04", "2011-03-06",                    "ADT",           "Bicalutamide",   "priorrx",
 "170-01-004", "2011-02-11", "2012-04-30",                    "ADT",             "Leuprolide",   "priorrx",
 "170-01-004", "2013-05-20", "2014-05-18",                    "ADT",             "Leuprolide",   "priorrx",
 "170-01-004", "2015-02-13", "2015-11-12",                    "ADT",             "Leuprolide",   "priorrx",
 "170-01-004", "2016-06-20", "2017-01-06",                    "ADT",           "Bicalutamide",   "priorrx",
 "170-01-004", "2016-06-22", "2017-09-07",                    "ADT",             "Leuprolide",   "priorrx",
 "170-01-004", "2006-12-15", "2007-06-15",                    "ADT",             "Leuprolide",   "priorrx",
 "170-01-004", "2018-04-18",           NA,                    "ADT",             "Leuprolide", "treatment",
 "170-01-004", "2018-04-18", "2019-01-30",                    "ADT",           "Bicalutamide", "treatment",
 "170-01-004", "2019-10-01", "2020-01-22", "AR Signaling Inhibitor",            "Apalutamide", "treatment",
 "170-01-004", "2020-02-11",           NA, "AR Signaling Inhibitor", "Darolutamide (ODM-201)", "treatment"
 ) |> 
  DT::datatable(
    options = list(pageLength = 20), 
    rownames = FALSE
    )

```



```{r, eval=FALSE, echo=TRUE}
# assign line of therapy 
treatment_lot <- treatment_prep %>% 
  # sort by subject and treatment start date
  arrange(subject, exstdat_int) %>%
  # within each subject, assign line of therapy
  group_by(subject) %>%
  group_modify(~ ironer::assign_lot(
    .data = .x, 
    # note, you could use treatment_category here as well to get a high level
    # view of line of therapy by category
    treatment = treatment_recode,
    dt_trt_observed = exstdat_int, 
    dt_treatment_end = exendat_int
    )) |> 
  ungroup()  
```


Example results from a subject. Note that individual treatments are still retained.

```{r, echo=FALSE, eval = TRUE}
tibble::tribble(
        ~subject, ~exstdat_int, ~exendat_int,      ~treatment_category,        ~treatment_recode,     ~source, ~lot,                   ~regimen, ~dt_lot_start, ~dt_lot_last_obs,
    "170-01-004", "2006-12-15", "2007-06-15",                    "ADT",           "Bicalutamide",   "priorrx",    1, "Bicalutamide, Leuprolide",  "2006-12-15",     "2007-06-15",
    "170-01-004", "2006-12-15", "2007-06-15",                    "ADT",             "Leuprolide",   "priorrx",    1, "Bicalutamide, Leuprolide",  "2006-12-15",     "2007-06-15",
    "170-01-004", "2011-02-04", "2011-03-06",                    "ADT",           "Bicalutamide",   "priorrx",    2, "Bicalutamide, Leuprolide",  "2011-02-04",     "2012-04-30",
    "170-01-004", "2011-02-11", "2012-04-30",                    "ADT",             "Leuprolide",   "priorrx",    2, "Bicalutamide, Leuprolide",  "2011-02-04",     "2012-04-30",
    "170-01-004", "2013-05-20", "2014-05-18",                    "ADT",             "Leuprolide",   "priorrx",    3,               "Leuprolide",  "2013-05-20",     "2015-11-12",
    "170-01-004", "2015-02-13", "2015-11-12",                    "ADT",             "Leuprolide",   "priorrx",    3,               "Leuprolide",  "2013-05-20",     "2015-11-12",
    "170-01-004", "2016-06-20", "2017-01-06",                    "ADT",           "Bicalutamide",   "priorrx",    4, "Bicalutamide, Leuprolide",  "2016-06-20",     "2017-09-07",
    "170-01-004", "2016-06-22", "2017-09-07",                    "ADT",             "Leuprolide",   "priorrx",    4, "Bicalutamide, Leuprolide",  "2016-06-20",     "2017-09-07",
    "170-01-004", "2018-04-18", "2019-01-30",                    "ADT",           "Bicalutamide", "treatment",    5, "Bicalutamide, Leuprolide",  "2018-04-18",     "2019-01-30",
    "170-01-004", "2018-04-18",           NA,                    "ADT",             "Leuprolide", "treatment",    5, "Bicalutamide, Leuprolide",  "2018-04-18",     "2019-01-30",
    "170-01-004", "2019-10-01", "2020-01-22", "AR Signaling Inhibitor",            "Apalutamide", "treatment",    6,              "Apalutamide",  "2019-10-01",     "2020-01-22",
    "170-01-004", "2020-02-11",           NA, "AR Signaling Inhibitor", "Darolutamide (ODM-201)", "treatment",    7,   "Darolutamide (ODM-201)",  "2020-02-11",     "2020-02-11"
    ) |> 
  DT::datatable(
    options = list(pageLength = 20), 
    rownames = FALSE
    )
```

You can reduce the data set to summarized treatment regimens.

```{r, eval=FALSE, echo=TRUE}
treatment_lot_summary <- treatment_lot |> 
  select(subject, lot, regimen, dt_lot_start, dt_lot_last_obs) |> 
  distinct()
```

And here is what the same subject would appear as:

```{r, echo=FALSE, eval = TRUE}
tibble::tribble(
      ~subject, ~lot,                   ~regimen, ~dt_lot_start, ~dt_lot_last_obs,
  "170-01-004",    1, "Bicalutamide, Leuprolide",  "2006-12-15",     "2007-06-15",
  "170-01-004",    2, "Bicalutamide, Leuprolide",  "2011-02-04",     "2012-04-30",
  "170-01-004",    3,               "Leuprolide",  "2013-05-20",     "2015-11-12",
  "170-01-004",    4, "Bicalutamide, Leuprolide",  "2016-06-20",     "2017-09-07",
  "170-01-004",    5, "Bicalutamide, Leuprolide",  "2018-04-18",     "2019-01-30",
  "170-01-004",    6,              "Apalutamide",  "2019-10-01",     "2020-01-22",
  "170-01-004",    7,   "Darolutamide (ODM-201)",  "2020-02-11",     "2020-02-11"
  )|> 
  DT::datatable(
    options = list(pageLength = 20), 
    rownames = FALSE
    )
```

